# Data Exploration

Before starting with data manipulation, the uswe begin with exploring the dataset.

## Data Structure

Let's have a look at the data structure of the Titanic dataset:

**R**
```{r}
# read the dataset
titanic <- read.csv("titanic.csv", stringsAsFactors = FALSE)
# give us the data stucture
str(titanic)
```

The output gives us a nice overview of the columns and its data structure, as well as a preview the underlying data. From this, we derive that the number of numeric columns is almost equal to the number of the character columns

**Python**
```{python}
import pandas as pd
# read the data set
titanic = pd.read_csv("titanic.csv")
# print the data structure
print titanic.dtypes
```

The Python ouptut is briefer than the R variant. Nevertheless, it gives a nice overview of the underlying structure. Interesting to note is that Python has given some columns other type definitions than R, such as "object", while in R the same columns have been labeled as character. The reason is that these columns contain different data types, causing Pandas to label it as "object".

## Data Summary

To get a bird eye view of the data, we can make a summary:

```{r}
titanic = read.csv("titanic.csv")
summary(titanic)
```
```{python}
import pandas as pd
titanic = pd.read_csv("titanic.csv")
print titanic.describe()
```



