--- 
title: "THE ULTIMATE GUIDE TO DATA MANIPULATION WITH R AND PYTHON"
author: "Zakaria Al Azhar, bigdatahabits.com"
date: "`r Sys.Date()`"
fontsize: 12pt
bibliography:
- book.bib
- packages.bib
description: This is a minimal example of using the bookdown package to write a book.
  The output format for this example is bookdown::gitbook.
documentclass: book
link-citations: yes
site: bookdown::bookdown_site
biblio-style: apalike
---

# Klm

<!--chapter:end:index.Rmd-->


# Introduction {#intro}

<!--chapter:end:01-intro.Rmd-->


# Prerequisites
## Basic knowledge
## Dataset
## Python and R packages

<!--chapter:end:02-Prerequisites.Rmd-->

# Data Exploration

Before starting with data manipulation or anything else, we begin with exploring the dataset.

## Data Structure

Let's have a look at the data structure of the Titanic dataset:

**R**
```{r}
# read the dataset
titanic <- read.csv("titanic.csv", stringsAsFactors = FALSE)
# give us the data stucture
str(titanic)
```

The output gives us a nice overview of the columns and its data structure,  even some preview of the underlying data. We can derive that the number of numeric columns is almost equal to the number of the character columns

**Python**
```{python}
import pandas as pd
# read the data set
titanic = pd.read_csv("titanic.csv")
# print the data structure
print titanic.dtypes
```

The Python ouptut is briefer than the R variant. Nevertheless, it gives a nice overview of the underlying structure. Interesting to note is that Python labels some columns as "object", while in R the same columns have been labeled as character. The reason is that these columns contain different data types, causing Pandas to label it as "object".

## Data Summary

To get a bird eye view of the data, we can make a summary:

```{r}
summary(titanic)
```
```{python}
import pandas as pd
titanic = pd.read_csv("titanic.csv")
print titanic.describe()
```




<!--chapter:end:03-Data_Exploration.Rmd-->


# Column Formulas
## Data Binning
## Convert & Replace
### Category to Number

<!--chapter:end:04-Column_Data_Manipulation.Rmd-->


# Final Words and this is strange

<!--chapter:end:05-summary.Rmd-->


# Placeholder

<!--chapter:end:06-references.Rmd-->

